A Non-local Bayesian Video Denoising Method

### 2016-07-20 doc/201507_tip/paper.tex

\begin{abstract}
	The quality provided by image/video sensors increases steadily, and
	for a fixed spatial resolution the sensor noise has been gradually reduced.
	However, modern sensors are also capable of acquiring at higher spatial
	resolutions which are still affected by noise, specially at low lighting
	conditions. The situation is even worse in video cameras, where the capture
	time is bounded by the frame rate. The noise in the video degrades its visual 
	quality and hinders its analysis.
	In this paper we present a new video denoising method extending 
	the non-local Bayes image denoising algorithm. The method does not require 
	motion estimation, and yet preliminary results show that it outperforms the
	state-of-the-art methods in terms of PSNR.
\end{abstract}

@@

Video denoising has received much less attention than still image denoising in
the literature.  In principle, denoising videos should be easier than still
images, since the strong temporal redundancy along motion trajectories favours
the denoising task. However, this additional form of redundancy creates
challenges.  On one hand, the amount of data is much larger than for still
images, and efficient ways of navigating through this data are needed. On the
other hand, the output of the denoising algorithm should be temporally
consistent according to the motion of the original sequence.

@@

The methods in \cite{Buades2005v} and \cite{Dabov2007v} are extensions to video
of still image patch-based denoising algorithms (non-local means \cite{Buades2006} and BM3D
\cite{Dabov2007tip} respectively). They are based on filtering similar 2D patches
searched for in a 3D spatio-temporal neighborhood. These methods exploit both
spatial and temporal redundancy, however, since each 2D patch is processed
independently, there is no mechanism to impose coherence along trajectories.
This results in flickering artifacts which become noticeable for high levels of
noise. 
%
While these methods exploit spatio-temporal redundancy, they do not impose
coherence of the trajectories, resulting in flickering
artifacts which become noticeable for high levels of noise.

@@

Motion-compensated patches have the benefit that their spatial slices are very
similar between each other. As a consequence, temporal filtering can be easily performed by
filtering inside the patch in the temporal direction. 
Because of this, methods that use motion-compensated 3D patches only look for
similar motion-compensated 3D patches starting in the same frame.

@@

One of the main difficulties in using motion-compensated patches is estimating
motion. Motion estimation is a paramount problem in noiseless sequences, and
noise makes it much harder.

Rectangular 3D patches offer an appealing
alternative, since they do not require estimating motion. 
In \cite{Protter2007,Protter2009} the authors extend the K-SVD \cite{Elad2006}
image denoising method to video by learning a dictionary of spatio-temporal
rectangular 3D patches.

Each rectangular 3D patch results from the combination of a 2D texture and a
motion pattern. For arbitrary motion, each combination of texture and motion is
not likely to occur together again. Thus rectangular 3D patches should have
less repeatability in the video compared to motion compensated patches.  An
exception is when the motion is a global translation of the frame, in which
case, the motion pattern is equal for each patch. Furthermore, if there is no
acceleration (uniform motion), rectangular 3D patches are constant along motion
trajectories. Global uniform motion is indeed a too restrictive model, but in
many cases, it is a valid approximation in a local neighborhood in space and
time.

@@

In this work we present new Bayesian patch-based video denoising method using
3D rectangular patches and not requiring motion estimation.
Yet, it provides
results which compare favourably to the current state-of-the-art,
both quantitatively by a significant margin and qualitatively, specifically in
terms of the temporal consistency attained.
In the spirit of  \cite{Lebrun2013a,Lebrun2013ipol},
the method works by learning a Gaussian \emph{a priori} model from sets of 
similar noisy patches. Each patch is denoised by computing the \emph{maximum a
posteriori} (MAP) with the estimated prior.
Learning the Gaussian prior amounts to estimating a covariance matrix with a
limited number of samples in a high dimensional space. Motivated by recent 
theoretical results and supported by an empirical analysis, we add the 
assumption that the \emph{a priori} model is of low rank (\ie the clean patches
lie in a lower dimensional subspace). This assumption lowers the computational cost
of the algorithm, and improves the results.

@@

The rest of the paper is organized as follows: in \S
\ref{sec:review_nonlocal_Bayes} we review our patch-based Bayesian optimal estimator for
video denoising. The final algorithm and details about its implementation are
given in \S \ref{sec:algorithm}.
Some relevant aspects of the proposed method are discussed in \S
\ref{sec:discussion}.
In \S \ref{sec:parameters}
we explain the selection of the parameters. Some results, including a
comparison with the state-of-the-art methods, are shown in \S
\ref{sec:results}. Concluding remarks are given in \S \ref{sec:conclusion}.

@@

The parameters of the \textit{a priori} model are learnt from the noisy video as
follows. For each noisy patch $\ma q$, we select the most $n$
similar 3D patches found in a search region centered at $\ma q$. The
search region is a spatio-temporal rectangle of size $w_x \times
w_x \times w_t$ ($w_x, w_t$ are odd numbers).

Let us note by $\ma q_i$, $i = 1, \dots, n$ the set of patches
similar to $\ma q$ (with $\ma q_1 = \ma q$). We assume that these 
patches correspond to different noisy observations of the Gaussian linear model
\eqref{eq:prior} and \eqref{eq:obs}. The corresponding marginal distribution is given by
\[\mathds P(\ma q) = \mathcal N(\overline{\ma p}, C + \sigma^2I). \]
Therefore, the maximum likelihood estimates for $\overline{\ma p}$ and $C$ are given by 
\begin{equation}
	\widehat{\overline{\ma p}} = \frac1{n}\sum_{i = 1}^{n}\ma q_i \quad\text{and}\quad 
		\widehat C= \frac1n\sum_{i = 1}^{n}\ma q_i\ma q_i^T - \sigma^2I.
	\label{eq:learn_parameters}
\end{equation}

We will also assume that the covariance matrix has a lower rank $k$.  In this
case, it has been shown \cite{Tipping1999} that the maximum likelihood estimate is given by
truncating the spectral decomposition of $\widehat C$ to the $k$ largest
eigenvalues 
\begin{equation}
\widehat C_k = \sum_{i = 1}^k\widehat \lambda_i\widehat U_i\widehat U_i^T,
\end{equation}
where $\widehat \lambda_i$, $\widehat U_i$ denote the $i$-th largest
eigenvalue and its corresponding eigenvector respectively.

\padd{The estimated eigenvalues are given by $\widehat \lambda_i = \widehat
\nu_i - \sigma^2$, where $\widehat \nu_i$ are the eigenvalues of the empirical
covariance matrix of the noisy patches. In practice, some of the smallest
estimated eigenvalues $\widehat \nu_i$ can be smaller than $\sigma^2,$ causing
$\widehat \lambda_i$ and the corresponding filter coefficient $w_{ii}$ to be
negative. Negative variances do not have physical meaning, and the
corresponding weights have a negative impact on the MAP estimate. This is an
issue of empirical Wiener filters and it is typically addressed by zeroing out
the negative coefficients. Later in \S
\ref{sec:parameters} we compare the results obtained by zeroing out negative
coefficients against using them.}

@@

Figure \ref{fig:patch_groups_patches} shows the 20 nearest neighbors in 
more detail. Each patch is displayed by vertically stacking its four $9\times
9$ spatial slices. For the two examples shown in Figure \ref{fig:patch_groups_pos}
the figure shows the noisy patches, the result of the MAP estimation and the
mean patch together with the leading 19 principal directions. The Gaussian
models were learnt from the $n = 200$ similar patches shown in Figure
\ref{fig:patch_groups_pos}. The MAP estimates were computed setting $r = 40$.
Since the patch is not
motion-compensated, all spatial slices can be different,
and thus the same applies to the principal directions.
The spatial slices of the principal directions show oscillating patterns, with
increasing frequency. In some cases these patterns are localized in a part of
the spatial slice.

@@

Other methods for image denoising have been proposed in the
literature based on Gaussian models for the patch density. In \cite{Yu2012},
the authors introduce a framework for solving inverse problems, based on the
assumption that the patches of an image are distributed according to a Gaussian
Mixture Model (GMM) which is learnt from
the image. A GMM is also used in \cite{Zoran2011}, but it is trained from a
database of $2\cdot 10^6$ patches randomly sampled from the Berkeley database.

In \cite{Zhang2010} a PCA-based algorithm is introduced which is equivalent to 
the nonlocal Bayes method. In fact, nonlocal Bayes can be regarded as a Bayesian 
interpretation of \cite{Zhang2010}.

The BLS-GSM method \cite{Portilla2003} (Bayes least square estimate of Gaussian
scale mixture) models noiseless ``wavelet coefficient neighborhoods'' with a
Gaussian scale mixture defined as a random scaling of a zero-mean Gaussian
density. The wavelet coefficient neighborhood turns out to be a patch of an
oriented channel of the image at a given scale.

\begin{figure*}[htpb!]
	\centering
	\includegraphics[width = .8\textwidth]{figs/patch_groups/patch_group_bus_045_085_012_s40_wx37_wt2_sx9_st4_r040_n200_coor.png}\\
	\vspace{.2cm}
	\includegraphics[width = .8\textwidth]{figs/patch_groups/patch_group_bus_255_056_010_s40_wx37_wt2_sx9_st4_r040_n200_coor.png}
%	\includegraphics[width = .8\textwidth]{figs/patch_groups/patch_group_bus_110_090_010_s40_wx37_wt2_sx9_st4_r040_n200_coor.png}
	\caption{Examples of sets of 200 nearest neighbors spatio-temporal patches,
	for patches of size $9\times 9\times 4$.
	The images in each row show the five frames of a $37\times37\times5$ search region. 
	In red we show the positions of the 5 nearest neighbors, in green the next 40 and in blue the rest.
	Note that the points shown correspond to the top-left pixel of the first spatial
	slice of each patch.
	The reference patch is the red patch in the center of the search region. To highlight the
	position of the patches, the color of the images has been attenuated.}
	\label{fig:patch_groups_pos}
\end{figure*}

@@

Let us now describe the video denoising algorithm. As in
\cite{Dabov2007tip,Dabov2007v,Maggioni2012,Lebrun2013a} we perform two stages, described in 
Algorithms \ref{alg:nlbayes} and \ref{alg:nlbayes2}. In the first stage we
compute a \emph{basic} estimate, which we denote $\widetilde u^{(1)}$.

In the second stage computes the final estimate $\widetilde u^{(2)}$ using the
basic estimate as an oracle. This entails two modifications with respect to the
first stage (highlighted in italics in Algorithm \ref{alg:nlbayes2}). Firstly, the
computation of the patch distances is performed using the patches of the basic
estimate instead of the noisy ones. And secondly, we assume that the patches of
the basic estimate are drawn from the \textit{a priori} Gaussian model and use
them to learn the mean and covariance matrix. Note that in the second stage,
when estimating the covariance matrix, we assume that the patches of the basic
estimate have no noise. Thus we set $\sigma = 0$ in Eq.
\eqref{eq:learn_parameters}.

@@

\paragraph{Patch distance threshold} When building the group of similar
patches during the second stage, more than $n$ patches are allowed
if their distances with respect to the reference patch are smaller than a
threshold $\tau$.  As in \cite{Lebrun2013ipol}, we set $\tau = 4$ (the distance between
patches is normalized by the number of elements in the patch and the number of
channels and the values of each channel are in the range  $[0,255]$).

@@

The computation can be substantially accelerated by reducing the 
number of groups of patches that are jointly processed.
Each group of patches is associated to the reference patch from which 
the nearest neighbors are computed.
We apply two tricks
considered in \cite{Dabov2007tip,Lebrun2013ipol} in order to reduce the number of reference patches. 
The first one consists on visiting only 
those patches on a coarser spatial grid with vertical and horizontal separation
of $s_x/2$. This reduces the number of reference patches by a factor of
$s_x^2/4$. The second trick consists on the following.  After
filtering the patches in a group, these patches are not considered again
as reference patches. 
Note that we ensure that each reference patch is filtered at least once. 
Due to the spacing of the coarse grid, this guarantees that each pixel is
overlapped by four reference patches on the same frame, so it will receive at
least four contributions in the aggregation. The actual number of aggregated
estimates for each pixel is much higher, considering the reference and
non-reference patches (which can also be filtered) in the neighboring frames.

Due to these speed-ups, the actual number of groups of similar patches processed
depends on $n$, and varies between a $1\%$ ($n \approx 400$)  and a $10\%$ ($n
\approx 50$) of the total number of pixels in the sequence. 
\pcomment{Check this numbers!}

@@

The rank also plays a role in the estimation of the covariance matrix.
As explained in \S \ref{sse:complexity}, alleviating the computational cost is
one of the reasons for using a low rank approximation of the empirical
covariance matrix. But there are also good reasons regarding the denoising
performance. In situations in which the number of samples 
is comparable to their dimension, the spectral decomposition of the empirical
covariance matrix $\widehat C$ can have considerable errors with respect to
that of the population covariance matrix $C$. 
%
The case of Gaussian models with covariance matrices of the form $C = A + \sigma^2I$,
where $A$ does not have full rank, has been studied in the statistics and physics
literature. The authors in \cite{johnstone2009,paul2007,nadler2008} study the
limit when the number of samples $n$ and the dimension $p$ grow to infinity at
same rate, such that $p/n\to\gamma < \infty$.  These works show that, for a
non-zero ratio $\gamma$, the empirical eigenvectors 
and eigenvalues are inconsistent estimators of the population ones.
Furthermore, the smaller eigenvalues and their eigenvectors will have a higher
error and eventually, if the signal-to-noise ratio drops bellow $\sqrt{\gamma}$, 
the eigenvector is confused with the noise and the resulting estimate will be,
in expectation, orthogonal to the actual value.
Finite sample bounds for the eigenvalues and eigenvectors have been also studied 
by \cite{nadler2008}, in the case in which $A$ has rank 1.
\pcomment{What can we say about this bounds.}

The errors between the sample eigenvectors and their population
counterparts manifest themselves as noise, which then affects the MAP estimate.
%
These asymptotic results suggest to discard the eigenvectors associated with 
lower eigenvalues. How many should be discarded in practice when $n$ and $p$ are finite, 
can only be fixed empirically, as we do in \$ \ref{sec:parameters}.

@@

The number of similar patches $n$ and the rank $r$ affect the estimation of the covariance
matrix $C$, controlling the trade-off between the accuracy of the Gaussian model
and the precision of the covariance matrix,
which in turn has a direct effect on the denoising strength of the algorithm.
% This have great impact on the bias-variance trade-off of the denoiser.
%In practice, these parameters determine the denoising strength of the
%algorithm. For example, using a large
We found a compromise to this trade-off empirically, by trying different values
of this parameter on a few training sequences and keeping the one that yields
the best PSNR on average.

@@

\padd{For the basic estimate, we show two set of curves. The dashed curves were
obtained by zeroing out negative eigenvalues of the estimated a priori covariance
matrix $\widehat C$, computed according to \eqref{eq:learn_parameters}. The continuous curves 
were obtained by retaining negative eigenvalues. In this case the MAP is estimated
with an empirical Wiener filter that may have some negative coefficients.}

\padd{For lower values of the rank parameter $r$, both curves coincide, since the first
eigenvalues are positive. 
For all noise levels, all curves of fixed $n_1$ present a first peak at a low value of the rank $r$.
The behaviour of the basic estimate around this peak can be
interpreted in the light of the theoretical results described in \S
\ref{sse:covariance-estimation}.
As the noise level increases, the rank at which the peak is attained
shifts to the left. This makes sense, as the reliability of the computed eigenvectors
decrease with increasing noise.
%
Also, for a fixed noise level, when the number of samples $n_1$ increases, the value
of the peak increases also, and so does the rank at which it is attained.
This is again explainable, as smaller groups obtained by keeping only the
most similar patches have less modes of variation, and thus a worse signal-to-noise ratio
in the eigenvectors with lower eigenvalues.}
%As the group of patches grows
%bigger, patches in the group are not as similar anymore, and a higher rank is
%needed to allow for additional modes of variation.

%The eigenvalues of $\widehat C$ are given by $\hat \lambda_i = \widehat \nu_i -
%\sigma^2$, where $\widehat \nu_i$ are the eigenvalues of the empirical
%\emph{a posteriori} covariance matrix.
%In practice due to the estimation errors, several
%$\widehat \nu_i$ are smaller than $\sigma^2$. These translate into negative
%coefficients $w_{ii}$ in the empirical Wiener filter \eqref{eq:wiener_coeffs},
%which result in a poorer estimation of the MAP.
\padd{As the rank parameter increases, both curves diverge. 
The dashed curves become almost constant at the point when
most eigenvalues become negative. After that point increasing the
rank has no effect if the negative eigenvalues are zeroed out.
%
Surprisingly, these negative coefficients have a positive effect on the final 
result, creating the second peak for high values of the rank shown by the
continuous plots. Currently we do not have an explanation for this behaviour
although it is related to the aggregation: even if the estimated MAP for the
patches in each group have larger errors, after aggregation these errors seem
to cancel, yielding an overall better estimate of the clean image.}

\padd{The final estimate shows the same qualitative behaviour with respect to the first
stage parameters. The dashed curves obtained by zeroing out the negative
eigenvalues are not shown to avoid cluttering the figure. Essentially the
second step has the effect of reducing the differences between results obtained
with different parameters. Based on these curves, we fix the first stage parameters at
$n_1 = 300$ and $r_1 = 95$ for color videos, and at $n_1 = 300$ and $r_1 = 180$ for grayscale videos.}

>>>>

\begin{abstract}
In this paper we present a new patch-based empirical Bayesian video denoising algorithm.
The method builds a Bayesian model for each group of similar space-time
patches. These patches are not motion compensated, and therefore avoid the
risk of inaccuracies caused by motion estimation errors. The main drawback of empirical Bayesian
methods is the computational load for estimating rich statistics in high
dimension. We therefore explore dimensionality reduction strategies on groups
of similar patches. Our analysis, based on the literature
on empirical covariance estimation, leads to a significant complexity
reduction and better denoising performance. A comparison on classic benchmark
videos demonstrates improved visual quality and an increased PSNR with respect
to state-of-the-art video denoising methods.
\end{abstract}

@@

Video denoising has received much less attention than still image denoising in
the literature.  In principle, denoising videos should be easier than still
images, since the strong temporal redundancy along motion trajectories favours
the denoising task. However, this additional form of redundancy creates
challenges.  On one hand, the amount of data is much larger than for still
images, and efficient ways of navigating through this data are needed. On the
other hand, the output of the denoising algorithm should be temporally
consistent with to the motion of the original sequence, which is an essential
quality criterion.

@@

The methods in \cite{Buades2005v} and \cite{Dabov2007v} are extensions to video
of still image patch-based denoising algorithms (non-local means \cite{Buades2006} and BM3D
\cite{Dabov2007tip} respectively). They are based on filtering similar 2D patches
searched for in a 3D spatio-temporal neighborhood. These methods exploit both
spatial and temporal redundancy, however, since each 2D patch is processed
independently, there is no mechanism to impose coherence along trajectories.
This results in flickering artifacts which become noticeable for high levels of
noise. 
%

@@

Motion-compensated patches have the benefit that their spatial slices are 
to between each other. As a consequence, temporal filtering can be easily performed by
filtering inside the patch in the temporal direction. 
Because of this, methods that use motion-compensated 3D patches only look for
similar motion-compensated 3D patches starting in the same frame.

@@

One of the main difficulties in using motion-compensated patches is motion
estimation in the presence of noise.

Rectangular 3D patches offer an appealing
alternative, since they do not require motion estimation.
In \cite{Protter2007,Protter2009} the authors extend the K-SVD \cite{Elad2006}
image denoising method to video by learning a dictionary of spatio-temporal
rectangular 3D patches.

Each rectangular 3D patch results from the combination of a 2D texture and a
motion pattern. For arbitrary motion it would seem that each combination of
texture and motion is unlikely to repeat across the video. Thus rectangular 3D
patches should have less repeatability in the video compared to motion
compensated patches. 
This objection is mitigated in two ways. First, if the motion is uniform on the
trajectory, rectangular 3D patches will be similar along the trajectory.
Second, if the motion pattern is spatially homogeneous in a region of the
image, then the image self-similarity still extends to space-time patches
starting at the same frame, regardless of the complexity of the motion.

@@

In this work we present a new Bayesian patch-based video denoising method using
3D rectangular patches and not requiring motion estimation.
Yet, it provides
results which compare favourably to the current state-of-the-art,
both quantitatively by a significant margin and qualitatively, specifically in
terms of temporal consistency.
In the spirit of  \cite{Lebrun2013a,Lebrun2013ipol},
the method works by learning a Gaussian \emph{a priori} model from sets of 
similar noisy patches. Each patch is denoised by computing the \emph{maximum a
posteriori} (MAP) with the estimated prior.
The main difficulty of such a method is the problem of estimation a high
dimensional covariance matrix with a limited number of samples. 
%
There are two possible solutions to this problem. One that works in image
denoising, is to estimate a fixed \cite{Zoran2011} or an image dependent
Gaussian mixture model \cite{Chatterjee2012}.  The computational burden of the first
strategy is huge, but it is paid once, as it delivers to a fixed prior. And it
shows excellent performance. Yet its extension to video is challenging because
of the curse of dimensionality. The second strategy, of estimating a different
Gaussian mixture from each video, also raises serious complexity issues. Thus
we shall stick to the idea of using video redundancy to build a local Gaussian
model for each patch.
%
Motivated by recent 
theoretical results and supported by an empirical analysis, we add the 
assumption that the \emph{a priori} local patch model is of low rank (\ie the clean patches
lie in a lower dimensional subspace). This assumption lowers the computational cost
of the algorithm, and improves the results.

@@

The rest of the paper is organized as follows: in \S
\ref{sec:review_nonlocal_Bayes} we describe the proposed patch-based Bayesian optimal estimator for
video denoising. The resulting algorithm and details about its implementation are
given in \S \ref{sec:algorithm}.
The structure of spatio-temporal  patches and the issues of the estimation of
their covariance are discussed in \ref{sec:discussion}.
In \S \ref{sec:parameters}
we explain the selection of the parameters. Some results, including a
comparison with the state-of-the-art methods, are shown in \S
\ref{sec:results}. Concluding remarks are given in \S \ref{sec:conclusion}.

@@

The parameters of the \textit{a priori} model are learnt from the noisy video as
follows. For each noisy patch $\ma q$, we select the $n$ most
similar 3D patches found in a search region centered at $\ma q$. The
search region is a spatio-temporal rectangle of size $w_x \times
w_x \times w_t$ ($w_x, w_t$ are odd numbers).

Let us denote by $\ma q_i$, $i = 1, \dots, n$ the set of patches
similar to $\ma q$ (with $\ma q_1 = \ma q$). We assume that these 
patches correspond to noisy observations of the Gaussian linear model
\eqref{eq:prior} and \eqref{eq:obs}. The corresponding marginal distribution is given by
\[\mathds P(\ma q) = \mathcal N(\overline{\ma p}, C + \sigma^2I). \]
Therefore, the maximum likelihood estimates for $\overline{\ma p}$ and $C$ are given by 
\begin{equation}
	\widehat{\overline{\ma p}} = \frac1{n}\sum_{i = 1}^{n}\ma q_i \quad\text{and}\quad 
		\widehat C= \frac1n\sum_{i = 1}^{n}\ma q_i\ma q_i^T - \sigma^2I.
	\label{eq:learn_parameters}
\end{equation}

We will also assume that the covariance matrix has a lower rank $r$.  In this
case, it has been shown \cite{Tipping1999} that the maximum likelihood estimate is given by
truncating the spectral decomposition of $\widehat C$ to the $r$ largest
eigenvalues 
\begin{equation}
\widehat C_r = \sum_{i = 1}^r\widehat \lambda_i\widehat U_i\widehat U_i^T,
\end{equation}
where $\widehat \lambda_i$, $\widehat U_i$ denote the $i$-th largest
eigenvalue and its corresponding eigenvector respectively.

The estimated eigenvalues are given by $\widehat \lambda_i = \widehat
\nu_i - \sigma^2$, where $\widehat \nu_i$ are the eigenvalues of the empirical
covariance matrix of the noisy patches. In practice, some of the smallest
estimated eigenvalues $\widehat \nu_i$ can be smaller than $\sigma^2,$ causing
$\widehat \lambda_i$ and the corresponding filter coefficient $w_{ii}$ to be
negative. Negative variances do not have physical meaning, and the
corresponding weights have a negative impact on the MAP estimate. This is an
issue of empirical Wiener filters and it is typically addressed by zeroing out
the negative coefficients. Later in \S
\ref{sec:parameters} we compare the results obtained by zeroing out negative
coefficients against using them.

@@

Figure \ref{fig:patch_groups_patches} shows the 20 nearest neighbors in 
more detail. Each patch is displayed by vertically stacking its four $9\times
9$ spatial slices. For the two examples shown in Figure \ref{fig:patch_groups_pos}
the figure shows the noisy patches, the result of the MAP estimation and the
mean patch together with the leading 19 principal directions. The Gaussian
models were learnt from the $n = 200$ similar patches shown in Figure
\ref{fig:patch_groups_pos}. The MAP estimates were computed setting the rank $r = 40$.
%Since the patch is not
%motion-compensated, its spatial slices can have little correlation.
%The same applies to the spatial slices of the principal directions.
The spatial slices of the principal directions show oscillating patterns, with
increasing frequency. 
In some cases these patterns are localized in a smaller region.
%
These oscillations are oriented in the direction of the
dominant edges in the corresponding slices of the patches in the group.

@@

Using a Gaussian patch model for for image denoising has several antecedents.
In \cite{Yu2012},
the authors introduce a framework for solving inverse problems, based on the
assumption that the patches of an image are distributed according to a Gaussian
Mixture Model (GMM) which is learnt from
the image. A GMM is also used in \cite{Zoran2011}, but it is trained from a
database of $2\cdot 10^6$ patches randomly sampled from the Berkeley database.
In \cite{Zhang2010} an algorithm that applies PCA on groups of similar patches
is introduced. It can be interpreted as a non-local empirical Bayesian method.
%
The BLS-GSM method \cite{Portilla2003} (Bayes least square estimate of Gaussian
scale mixture) models noiseless ``wavelet coefficient neighborhoods'' with a
Gaussian scale mixture defined as a random scaling of a zero-mean Gaussian
density. The wavelet coefficient neighborhood turns out to be a patch of an
oriented channel of the image at a given scale.

\begin{figure*}[htpb!]
	\centering
	\includegraphics[width = .8\textwidth]{figs/patch_groups/patch_group_bus_045_085_012_s40_wx37_wt2_sx9_st4_r040_n200_coor.png}\\
	\vspace{.2cm}
	\includegraphics[width = .8\textwidth]{figs/patch_groups/patch_group_bus_255_056_010_s40_wx37_wt2_sx9_st4_r040_n200_coor.png}
%	\includegraphics[width = .8\textwidth]{figs/patch_groups/patch_group_bus_110_090_010_s40_wx37_wt2_sx9_st4_r040_n200_coor.png}
	\caption{Examples of sets of 200 nearest neighbors spatio-temporal patches,
	for patches of size $9\times 9\times 4$.
	The images in each row show the five frames of a $37\times37\times5$ search region. 
	In red we show the positions of the 5 nearest neighbors, in green the next 40 and in blue the rest.
	The points shown correspond to the top-left pixel of the first spatial
	slice of each patch.
	The reference patch is the red patch in the center of the search region. To highlight the
	position of the patches, the color of the images has been attenuated.}
	\label{fig:patch_groups_pos}
\end{figure*}

@@

Let us now describe the video denoising algorithm. As in
\cite{Dabov2007tip,Dabov2007v,Maggioni2012,Lebrun2013a} we perform two stages, described in 
Algorithms \ref{alg:nlbayes} and \ref{alg:nlbayes2}. The first stage 
computes a \emph{basic} estimate, which we denote $\widetilde u^{(1)}$.

The second stage computes the final estimate $\widetilde u^{(2)}$ using the
basic estimate as an oracle. This entails two modifications with respect to the
first stage (highlighted in italics in Algorithm \ref{alg:nlbayes2}). Firstly, the
computation of the patch distances is performed using the patches of the basic
estimate instead of the noisy ones. And secondly, we assume that the patches of
the basic estimate are drawn from the \textit{a priori} Gaussian model and use
them to learn the mean and covariance matrix. In the second stage,
when estimating the covariance matrix, we assume that the patches of the basic
estimate have no noise. Thus we set $\sigma = 0$ in Eq.
\eqref{eq:learn_parameters}.

@@

\paragraph{Patch distance threshold} When building the group of similar
patches during the second stage, more than $n$ patches are allowed
if their distances with respect to the reference patch are smaller than a
threshold $\tau$.  As in \cite{Lebrun2013ipol}, we set $\tau = 4$ (the distance between
patches is jointly normalized by the number of elements in the patch, the number of
channels, and the values of each channel are in the range  $[0,255]$).

@@

The computation can be substantially accelerated by reducing the 
number of groups of patches that are jointly processed.
Each group of patches is associated to the reference patch from which 
the nearest neighbors are computed.
We apply two tricks
considered in \cite{Dabov2007tip,Lebrun2013ipol} in order to reduce the number of reference patches. 
The first one consists in visiting only 
those patches on a coarser spatial grid with vertical and horizontal separation
of $s_x/2$. This reduces the number of reference patches by a factor of
$s_x^2/4$. The second acceleration trick is to give up taking as reference
patch the patches that have been already once filtered in another group. 
%Note that we ensure that each reference patch is filtered at least once. 
%Due to the spacing of the coarse grid, this guarantees that each pixel is
%overlapped by four reference patches on the same frame, so it will receive at
%least four contributions in the aggregation. The actual number of aggregated
%estimates for each pixel is much higher, considering the reference and
%non-reference patches (which can also be filtered) in the neighboring frames.
The actual number of aggregated estimates for each pixel remains anyway high,
simply because each pixel is contained in 100 different patches, and at least
25 of them have been filtered, either as reference patches, or as patches
similar to a reference patch.

Due to these speed-ups, the actual number of groups of similar patches processed
depends on $n$, and varies across video sequences. We have measured empirically
that it ranges roughly from $1\%$ ($n \approx 400$) to $10\%$ ($n \approx 50$) of the total
number of pixels in the sequence. 
\pcomment{Check this numbers!}

@@

The rank also plays a role in the estimation of the covariance matrix.
As explained in \S \ref{sse:complexity}, alleviating the computational cost is
one of the reasons for using a low rank approximation of the empirical
covariance matrix. But there are also good reasons regarding the denoising
performance. In situations in which the number of samples 
is comparable to their dimension, the spectral decomposition of the empirical
covariance matrix $\widehat C$ can have considerable errors with respect to
that of the population covariance matrix $C$. 
%
The case of Gaussian models with covariance matrices of the form $C = A + \sigma^2I$,
where $A$ does not have full rank, has been studied in the statistics and physics
literature. The authors in \cite{johnstone2009,paul2007,nadler2008} study the
limit when the number of samples $n$ and the dimension $p$ grow to infinity at
the same rate, $p/n\to\gamma < \infty$.  These works show that, for a
non-zero ratio $\gamma$, the empirical eigenvectors 
and eigenvalues are inconsistent estimators of the population ones.
Furthermore, the smaller eigenvalues and their eigenvectors will have a higher
error and eventually, if the signal-to-noise ratio drops below $\sqrt{\gamma}$, 
the eigenvector is confused with the noise and the resulting estimate will be,
in expectation, orthogonal to the actual value.
Finite sample bounds for the eigenvalues and eigenvectors have been also studied 
by \cite{nadler2008}, in the case in which $A$ has rank 1.
\pcomment{What can we say about this bounds.}

The errors between the sample eigenvectors and their population
counterparts manifest themselves as noise, which then affects the MAP estimate.
%
These asymptotic results suggest to discard the eigenvectors associated with 
lower eigenvalues. How many should be discarded in practice when $n$ and $p$ are finite, 
can only be fixed empirically, as we do in \S \ref{sec:parameters}.

@@

The number $n$ of similar patches and the rank $r$ affect the estimation of the covariance
matrix $C$, controlling the trade-off between the accuracy of the Gaussian model
and the precision of the covariance matrix,
which in turn has a direct effect on the denoising strength of the algorithm.
% This have great impact on the bias-variance trade-off of the denoiser.
%In practice, these parameters determine the denoising strength of the
%algorithm. For example, using a large
We found a compromise to this trade-off empirically, by trying different values
of this parameter on a few training sequences and keeping the one that yields
the best PSNR on average.

@@

{For the basic estimate, we show two sets of curves. The dashed curves were
obtained by zeroing out negative eigenvalues of the estimated a priori covariance
matrix $\widehat C$, computed according to \eqref{eq:learn_parameters}. The continuous curves 
were obtained by retaining negative eigenvalues. In this case the MAP is estimated
with an empirical Wiener filter that may have some negative coefficients.}

{For lower values of the rank parameter $r$, both curves coincide, since the first
eigenvalues are positive. 
For all noise levels, all curves with fixed $n_1$ present a first peak at a low value of the rank $r$.
The behaviour of the basic estimate around this peak can be
interpreted in the light of the theoretical results described in \S
\ref{sse:covariance-estimation}.
As the noise level increases, the rank at which the peak is attained
shifts to the left. This makes sense, as the reliability of the computed eigenvectors
decreases with increasing noise.
%
Also, for a fixed noise level, when the number of samples $n_1$ increases, the value
of the peak increases also, and so does the rank at which it is attained.
This is again explainable, as smaller groups obtained by keeping only the
most similar patches have fewer modes of variation, and thus a worse signal-to-noise ratio
in the eigenvectors with lower eigenvalues.}
%As the group of patches grows
%bigger, patches in the group are not as similar anymore, and a higher rank is
%needed to allow for additional modes of variation.

%The eigenvalues of $\widehat C$ are given by $\hat \lambda_i = \widehat \nu_i -
%\sigma^2$, where $\widehat \nu_i$ are the eigenvalues of the empirical
%\emph{a posteriori} covariance matrix.
%In practice due to the estimation errors, several
%$\widehat \nu_i$ are smaller than $\sigma^2$. These translate into negative
%coefficients $w_{ii}$ in the empirical Wiener filter \eqref{eq:wiener_coeffs},
%which result in a poorer estimation of the MAP.
{As the rank parameter increases, both curves diverge. 
The dashed curves become almost constant at the point when
most eigenvalues become negative. After that point increasing the
rank has no effect if the negative eigenvalues are zeroed out.
%
Surprisingly, these negative coefficients have a positive effect on the final 
result, creating the second peak for high values of the rank shown by the
continuous plots.
%Currently we do not have an explanation for this behaviour
%although it is related to the aggregation:
Even if the estimated MAP for the
patches in each group have larger errors, after aggregation these errors seem
to cancel, yielding an overall better estimate of the clean image.}

{The final estimate shows the same qualitative behaviour with respect to the first
stage parameters. The dashed curves obtained by zeroing out the negative
eigenvalues are not shown to avoid cluttering the figure. Essentially the
second step has the effect of reducing the differences between results obtained
with different parameters. Based on these curves, we fixed the first stage parameters at
$n_1 = 300$ and $r_1 = 95$ for color videos, and at $n_1 = 300$ and $r_1 = 180$ for grayscale videos.}

Using these parameters for the first stage, we analyzed the impact when varying the second stage
parameters $n_2$ and $r_2$. The plots obtained are shown in Figure
\ref{fig:rank2-nsim2-fpsnr}.
%
Note that in this case, the \emph{a
priori} covariance matrix is estimated from the basic estimate (without having
to subtract $\sigma^2I$). As a consequence, there are no negative filter coefficients.
%
In the second stage the quality drops when
using a number of patches $n_2$ higher than $100$. This differs from the first stage,
where a larger $n_1$ yielded a better PSNR.
The reason for this lies probably in the fact that the patches have less noise,
and a good estimate of the leading eigenvectors of the covariance matrix can be obtained
with less patches. Furthermore, smaller groups of patches
are better modelled by a Gaussian distribution.
%\pcomment{I don't have an explanation for the curves for $\sigma = 10$. Precisely
%in the lower noise level, the PSNR drops when $n_2$ is reduced, except for lower ranks.
%This contradicts the argument given: for less noise, it should be always better to consider
%less patches, not only for lower ranks.} 
Based on these plots, we set the following values for the parameters
\[r_2(\sigma) = 30 - \frac12(\sigma - 10),\quad\quad n_{2} = 100.\]
